---
title: "Rogers' Diffusion"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Rogers' Diffusion}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Read about Rogers' Diffusion of Innovations model in the [original publication](http://www.worldcat.org/oclc/1084564086) of 1962 or on [Wikipedia](https://en.wikipedia.org/wiki/Diffusion_of_innovations).

The version implemented here sets up 100 agents, all of which connect to 10 random others, and 10 are initially informed. Any informed agents inform all other agents that follow them.

```{r setup}
library(tidyabm)
library(dplyr)
library(ggplot2)
```

Model configuration parameters are combined here:

```{r configuration}
n_uninformed_agents <- 90
n_initially_informed_agents <- 10
```

Let's create the agents (or, their blueprints):

```{r agents}
agent_a <- create_agent() %>%
  set_characteristic(informed = FALSE,
                     connected = FALSE) %>%
  add_rule('initial connect',
           connected == FALSE,
           .consequence = function(me, abm) {
             for (i in 1:10) {
               random_agent <- get_random_agent(abm, me)
               me <- network_connect(me, random_agent)
             }
             me %>%
               set_characteristic(connected = TRUE,
                                  .overwrite = TRUE,
                                  .suppress_warnings = TRUE) %>%
               return()
          }) %>%
  add_rule('diffuse information/innovation',
           informed == TRUE,
           .consequence = function(me, abm) {
             me %>%
               network_spread('informed',
                              TRUE,
                              overwrite = TRUE,
                              suppress_warnings = TRUE) %>%
               return()
           })

agent_b <- agent_a %>%
  set_characteristic(informed = TRUE,
                     .overwrite = TRUE)
```

Here's the actual model (i.e., the environment):

```{r diffusion}
e <- create_network_environment(seed = 926742,
                                is_directed = TRUE) %>%
  add_agents(agent_a,
             n = n_uninformed_agents) %>%
  add_agents(agent_b,
             n = n_initially_informed_agents) %>%
  add_variable(share_informed = function(me, abm) {
    abm %>%
      convert_agents_to_tibble() %>%
      summarise(share_informed = sum(informed)/n()) %>%
      pull(share_informed) %>%
      return()
  }) %>%
  add_rule('stop when all are informed',
           share_informed == 1,
           .consequence = stop_abm) %>%
  init()
```

Finally, iteration. This is the step that takes a while. Also, as we are not sure whether it terminates, we go with a maximum of 20 iterations for now:

```{r iteration, out.width = '100%'}
e <- e %>% 
  iterate(max_iterations = 20)
```

It seemed to have stalled. Most likely, that's because of the initial setup of the connections. Let's take a look at it visually.

```{r graph_visualization}
e %>% 
  visualize()
```

Yes, there is no connection between all of the agents. To manually inspect how everything went, we can just look at the environment or have it output all the agents.

```{r inspection}
e

e %>% 
  convert_agents_to_tibble()
```

Ultimately, we are interested in how fast innovation went:

```{r visualizations, out.width = '100%'}
e %>% 
  ggplot(aes(x = .tick,
             y = 100*share_informed)) +
  geom_line() +
  scale_x_continuous('Time [ticks]') +
  scale_y_continuous('Share of informed agents [%]') +
  theme_minimal() +
  ggtitle('Information/Innovation Diffusion over Time')
```

We can compile relevant data for any publication we're preparing through the [ODD protocol](https://doi.org/10.18564/jasss.4259):
```{r odd}
e %>% 
  odd()
```
