---
title: "Schelling's Segregation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Schelling's Segregation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Read about Schelling's Segregation model in the [original publication](https://doi.org/10.1080%2F0022250x.1971.9989794) of 1971 or on [Wikipedia](https://en.wikipedia.org/wiki/Schelling%27s_model_of_segregation).

The version implemented here sets up a squared grid of 20 by 20 spots. Neighborhood is defined as the surrounding up-to 8 fields of a particular agent. 

```{r setup}
library(tidyabm)
library(dplyr)
library(ggplot2)
```

Model configuration parameters are combined here:

```{r configuration}
grid_side_length <- 20
density <- .70
neighborhood_definition <- 'o'
unhappy_similarity_threshold <- .15
```

Let's create the agents (or, their blueprints):

```{r agents}
agent_a <- create_agent() %>%
  set_characteristic(color = 'red') %>%
  add_variable(similar = function(me, abm) {
    neighbors <- grid_get_neighbors(me, 
                                    abm, 
                                    which = neighborhood_definition)
    n_neighbors <- nrow(neighbors)
    if (n_neighbors == 0) {
      return(0.0)
    } else {
      return(sum(neighbors$color == me$color)/n_neighbors)
    }
  }) %>%
  add_variable(unhappy = function(me, abm) {
    return(me$similar < unhappy_similarity_threshold)
  }) %>%
  add_rule('move',
           unhappy == TRUE,
           .consequence = function(me, abm) {
             spot <- grid_get_free_spots(abm) %>% 
               slice_sample(n = 1)
             grid_move(me, 
                       abm,
                       new_x = spot$x,
                       new_y = spot$y) %>%
               return()
           })

agent_b <- agent_a %>%
  set_characteristic(color = 'blue',
                     .overwrite = TRUE)
```

Here's the actual model (i.e., the environment):

```{r segregation}
e <- create_grid_environment(seed = 38468,
                             size = grid_side_length) %>%
  add_agents(agent_a,
             n = grid_side_length^2 * density/2) %>%
  add_agents(agent_b,
             n = grid_side_length^2 * density/2) %>%
  add_variable(mean_similar = function(me, 
                                       abm) {
    abm %>%
      convert_agents_to_tibble() %>%
      summarise(M = mean(similar, na.rm = TRUE)) %>%
      pull(M) %>%
      return()
  }) %>%
  add_variable(share_unhappy = function(me, 
                                        abm) {
    abm %>%
      convert_agents_to_tibble() %>%
      summarise(share_unhappy = sum(unhappy)/dplyr::n()) %>%
      pull(share_unhappy) %>%
      return()
  }) %>%
  add_rule('stop when all are happy',
           share_unhappy <= 0,
           .consequence = stop_abm) %>%
  init()
```

Finally, iteration. This is the step that takes a while (but we can watch):

```{r iteration, out.width = '100%'}
e <- e %>% 
  iterate(max_iterations = 50,
          visualize = TRUE,
          color = color,
          shape = unhappy)
```

To manually inspect how everything went, we can just look at the environment or have it output all the agents.

```{r inspection}
e

e %>% 
  convert_agents_to_tibble()
```

Ultimately, we are interested in how average similarity and the share of unhappy agents went down:

```{r visualizations, out.width = '100%'}
e %>% 
  ggplot(aes(x = .tick,
             y = 100*share_unhappy)) +
  geom_line() +
  scale_x_continuous('Time [ticks]') +
  scale_y_continuous('Share of unhappy agents [%]') +
  theme_minimal() +
  ggtitle('Unhappiness over Time')


e %>% 
  ggplot(aes(x = .tick,
             y = 100*mean_similar)) +
  geom_line() +
  scale_x_continuous('Time [ticks]') +
  scale_y_continuous('Average share of similar agents in neighborhood [%]') +
  theme_minimal() +
  ggtitle('Mean Similarity over Time')
```

We can compile relevant data for any publication we're preparing through the [ODD protocol](https://doi.org/10.18564/jasss.4259):
```{r odd}
e %>% 
  odd()
```
